---
title: "Lab 0 - Mike G"
author: "Munkhnaran Gankhuyag"
date: "August 30, 2017"
output:
  pdf_document: default
  html_document: default
  word_document: default
---



##Questions    

```{r load-abrbuthnot-data, eval=TRUE}
source("more/arbuthnot.R")
```

1.  What command would you use to extract just the counts of girls baptized? Try
    it!
```{r}
arbuthnot$girls
```

2.  Is there an apparent trend in the number of girls baptized over the years?  
    How would you describe it?

```{r}
plot(x = arbuthnot$year, y = arbuthnot$girls, type = "l")
```

There is an increasing trend after a slight dip in 1650.

3.  Now, make a plot of the proportion of boys over time. What do you see? Tip: 
If you use the up and down arrow keys, you can scroll through your previous 
commands, your so-called command history. You can also access it by clicking on 
the history tab in the upper right panel.  This will save you a lot of typing in
the future.

```{r}
plot(x = arbuthnot$year, y = arbuthnot$boys, type = "l")

```
There is a similar trend among boys with slightly larger number of baptizes.



```{r load-present-data, eval=TRUE}
source("more/present.R")
```


#Part 1
```{r}
View(present)#shows years 1940 - 2002
dim(present) #63 observations of 3 variables
names(present) #"year" "boys" "girls"
```

#Part 2
```{r}
plot(x=present$year, y=present$girls +present$boys, type = "l")

summary(present)
```

Present data shows a boom in births during the mid 1940s, a slight dip during the 1980s and an upward tred after 1980. The Arbuthnot data showed the opposite. A downward trend in the 1650's and a boom after the 1660s. 

```{r}
plot(x=present$year, y=present$boys, type = "l")
summary(present$boys)
```

```{r}
plot(x=present$year, y=present$girls, type = "l")
summary(present$girls)
```

```{r}
max(present$boys + present$girls) -min(present$boys + present$girls)
```

While the graph shows similar trends, when you compare the actual numbers, the present number of births is significantly larger.
The range for the arbuthnot is 10533 births.
The range for the present is 1907927 births.


#Part 3
```{r}
plot(x= present$year, y =present$boys/present$girls, type = "l")
summary(present$boys/present$girls)
```

boys to girls ratio is all above 1, which means Arbuthnot's observation about boys being born in greater proportion than girls hold up in the U.S.

```{r}
present$boys>present$girls

```

When you test if present boys is greater than present girls, it shows 63 TRUE values.

#Part 4

```{r}
which.max(present$boys + present$girls)
present[22, ] 
```

1961 had the largest number of births.
